### 2016-06-29 2016_Proceedings_ISMIR/keynotes.tex
initial checkin, made some changes from 2015 already

%-------------------------------------------
\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Keynote Talks}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/10_Keynotes.pdf}
\thispagestyle{empty}\cleardoublepage
 %-------------------------------------------
% Keynote talk 1
\addcontentsline{toc}{subsection}{Integrating Music Information Sources for Music Production and Consumption\\\textit{Mark Sandler}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/10-1_Keynotes-1.pdf}%

\thispagestyle{empty}\cleardoublepage
% Keynote talk 2
\addcontentsline{toc}{subsection}{The Promise of Music Information Retrieval: Are we there yet?\\\textit{J.\ Stephen Downie}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/10-2_Keynotes-2.pdf}%