### 2016-06-29 2016_Proceedings_ISMIR/tutorials.tex

%-------------------------------------------
\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Tutorials}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/11_Tutorials.pdf}%
\thispagestyle{empty}\cleardoublepage
 %-------------------------------------------
% Tutorial 1
\addcontentsline{toc}{subsection}{Why Singing is Interesting\\\textit{Simon Dixon, Masataka Goto, Matthias Mauch}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/11-1_Tutorial-1.pdf}%

% Tutorial 2
\addcontentsline{toc}{subsection}{Addressing the Music Information Needs of Musicologists\\\textit{Richard J.\ Lewis, Ben Fields, Tim Crawford}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/11-2_Tutorial-2.pdf}%

% Tutorial 3
\addcontentsline{toc}{subsection}{Markov Logic Networks for Music Analysis\\\textit{Helene Papadopoulos}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/11-3_Tutorial-3.pdf}%

% Tutorial 4
\addcontentsline{toc}{subsection}{COmputation and FLAmenco: Why Flamenco is Interesting for
MIR Research\\\textit{Emilia G{\'o}mez, Nadine Kroher, Jose Miguel D{\'i}az-B{\'a}{\~n}ez, Sergio Oramas, 
Joaqu{\'i}n Mora, Francisco G{\'o}mez-Mart{\'i}n}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/11-4_Tutorial-4.pdf}%

% Tutorial 5
\addcontentsline{toc}{subsection}{Using Correlation Analysis and Big Data to Identify and Predict Musical Behaviors\\\textit{Jeff C.\ Smith}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/11-5_Tutorial-5.pdf}%

% Tutorial 6
\addcontentsline{toc}{subsection}{Automatic Music Transcription\\\textit{Zhiyao Duan, Emmanouil Benetos}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/11-6_Tutorial-6.pdf}%