### 2016-06-29 2016_Proceedings_ISMIR/tutorials.tex

%-------------------------------------------
\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Tutorials}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/11_Tutorials.pdf}%
\thispagestyle{empty}\cleardoublepage
 %-------------------------------------------
% Tutorial 1
\includeabstract{Why Singing is Interesting}{Simon Dixon, Masataka Goto, Matthias Mauch}{external/11-1_Tutorial-1.pdf}

% Tutorial 2
\includeabstract{Addressing the Music Information Needs of Musicologists}{Richard J.\ Lewis, Ben Fields, Tim Crawford}{external/11-2_Tutorial-2.pdf}

% Tutorial 3
\includeabstract{Markov Logic Networks for Music Analysis}{Helene Papadopoulos}{external/11-3_Tutorial-3.pdf}

% Tutorial 4
\includeabstract{COmputation and FLAmenco: Why Flamenco is Interesting for
MIR Research}{Emilia G{\'o}mez, Nadine Kroher, Jose Miguel D{\'i}az-B{\'a}{\~n}ez, Sergio Oramas, Joaqu{\'i}n Mora, Francisco G{\'o}mez-Mart{\'i}n}{external/11-4_Tutorial-4.pdf}

% Tutorial 5
\includeabstract{Using Correlation Analysis and Big Data to Identify and Predict Musical Behaviors}{Jeff C.\ Smith}{external/11-5_Tutorial-5.pdf}

% Tutorial 6
\includeabstract{Automatic Music Transcription}{Zhiyao Duan, Emmanouil Benetos}{external/11-6_Tutorial-6.pdf}