### 2016-06-29 2016_Proceedings_ISMIR/keynotes.tex
updated TOC info for keynotes and tutorials

%-------------------------------------------
\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Keynote Talks}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/10_Keynotes.pdf}
\thispagestyle{empty}\cleardoublepage
 %-------------------------------------------
% Keynote talk 1
\includeabstract{Keynote 1}{Speaker 1}{external/10-1_Keynotes-1.pdf}

\thispagestyle{empty}\cleardoublepage
% Keynote talk 2
\includeabstract{Keynote 2}{Speaker 2}{external/10-2_Keynotes-2.pdf}