### 2016-06-29 2016_Proceedings_ISMIR/tutorials.tex

%-------------------------------------------
\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Tutorials}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/11_Tutorials.pdf}%
\thispagestyle{empty}\cleardoublepage
 %-------------------------------------------
% Tutorial 1
\includeabstract{Jazz solo analysis between music information
  retrieval, music psychology, and jazz research}{Jakob Abe{\ss}er,
  Klaus Frieler, Wolf-Georg Zaddach}{external/11-1_Tutorial-1.pdf}

% Tutorial 2
\includeabstract{Music Information Retrieval: Overview, Recent
  Developments and Future Challenges}{Emilia G\'omez, Markus Schedl,
  Xavier Serra, Xiao Hu}{external/11-2_Tutorial-2.pdf}

% Tutorial 3
\includeabstract{Why is studio production interesting?}{Emmanuel
  Deruty, Fran\c{c}ois Pachet}{external/11-3_Tutorial-3.pdf}

% Tutorial 4
\includeabstract{Introduction to EEG Decoding for Music Information Retrieval
Research}{Sebastian Stober, Blair Kaneshiro}{external/11-4_Tutorial-4.pdf}

% Tutorial 5
\includeabstract{Natural Language Processing for MIR}{Sergio Oramas,
  Luis Espinosa-Anke, Shuo Zhang, Horacio Saggion}{external/11-5_Tutorial-5.pdf}

% Tutorial 6
\includeabstract{Why Hip-Hop is interesting}{Jan Van Balen, Ethan
  Hein, Dan Brown}{external/11-6_Tutorial-6.pdf}