### 2016-06-30 2016_Proceedings_ISMIR/papers.tex
wrote script to generate papers.tex, might not have gotten unicode right yet

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 1}
\includepdf[pages=2,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Learning to Pinpoint Singing Voice from Weakly Labeled Examples}{Jan Schlüter}{articles/315_Paper}
\includepaper{SiMPle: Assessing Music Similarity Using Subsequences Joins}{"Diego Silva, Chin-Chia Michael Yeh, Gustavo Batista and Eamonn Keogh"}{articles/099_Paper}
\includepaper{Feature Learning for Chord Recognition: The Deep Chroma Extractor}{Filip Korzeniowski and Gerhard Widmer}{articles/178_Paper}
\includepaper{Brain Beats: Tempo Extraction from EEG Data}{"Sebastian Stober, Thomas Praetzlich and Meinard Mueller"}{articles/022_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 2}
\includepdf[pages=3,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{AUTOMATIC TAGGING USING DEEP CONVOLUTIONAL NEURAL NETWORKS}{"Keunwoo Choi, George Fazekas and Mark Sandler"}{articles/009_Paper}
\includepaper{Analyzing Measure Annotations for Western Classical Music Recordings}{"Christof Weiß, Vlora Arifi-Müller, Thomas Prätzlich, Rainer Kleinertz and Meinard Müller"}{articles/079_Paper}
\includepaper{Score-Informed Identification of Missing and Extra Notes in Piano Recordings}{"Sebastian Ewert, Siying Wang, Meinard Mueller and Mark Sandler"}{articles/123_Paper}
\includepaper{Joint Beat and Downbeat Tracking with Recurrent Neural Networks}{"Sebastian Böck, Florian Krebs and Gerhard Widmer"}{articles/186_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 3}
\includepdf[pages=5,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Tempo Estimation for Music Loops and a Simple Confidence Measure}{Frederic Font and Xavier Serra}{articles/195_Paper}
\includepaper{Using Priors To Improve Estimates of Music Structure}{Jordan Smith and Masataka Goto}{articles/135_Paper}
\includepaper{A Neural Greedy Model for Voice Separation in Symbolic Music}{Patrick Gray and Razvan Bunescu}{articles/296_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 4}
\includepdf[pages=6,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Towards Reading Sheet Music}{"Matthias Dorfer, Andreas Arzt and Gerhard Widmer"}{articles/027_Paper}
\includepaper{A plan for sustainable MIR evaluation}{"Brian McFee, Eric Humphrey and Julián Urbano"}{articles/257_Paper}
\includepaper{Go With the Flow: When Listeners Use Music as Technology}{"Andrew Demetriou, Martha Larson and Cynthia C. S. Liem"}{articles/068_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 5}
\includepdf[pages=7,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Automatic Melodic Reduction using a Supervised Probabilistic Context-Free Grammar}{Ryan Groves}{articles/274_Paper}
\includepaper{Singing Melody Extraction using Multi-Column Deep Neural Networks}{"Sangeun Kum, Changheun Oh and Juhan Nam"}{articles/119_Paper}
\includepaper{Instrumental idiom in the 16th century: patterns in arrangements of vocal music}{"David Lewis, Tim Crawford and Daniel Müllensiefen"}{articles/165_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 6}
\includepdf[pages=9,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Learning a feature space for world music style similarity}{"Maria Panteli, Emmanouil Benetos and Simon Dixon"}{articles/166_Paper}
\includepaper{Bayesian meter tracking on learned signal representations}{Andre Holzapfel and Thomas Grill}{articles/132_Paper}
\includepaper{The Sousta corpus: Beat-informed automatic transcription of traditional dance tunes}{Andre Holzapfel and Emmanouil Benetos}{articles/156_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 7}
\includepdf[pages=10,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{A LOOK AT THE CLOUD FROM BOTH SIDES NOW: AN ANALYSIS OF CLOUD MUSIC SERVICE USAGE}{"Jin Ha Lee, Yea-Seul Kim and Chris Hubbles"}{articles/087_Paper}
\includepaper{A Hybrid Gaussian-HMM-Deep Learning Approach For Automatic Chord Estimation With Very Large Vocabulary}{Junqi Deng and Yu-Kwong Kwok}{articles/058_Paper}
\includepaper{SYSTEMATIC EXPLORATION OF COMPUTATIONAL MUSIC STRUCTURE RESEARCH}{Oriol Nieto and Juan Pablo Bello}{articles/043_Paper}
\includepaper{Extracting Ground-Truth Information from MIDI Files: A MIDIfesto}{Colin Raffel and Daniel Ellis}{articles/105_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Poster session 1}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{TIME-DELAYED MELODY SURFACES FOR R_GA RECOGNITION}{"Sankalp Gulati, Joan Serrà, Kaustuv Kanti Ganguli, Sertan _entürk and Xavier Serra"}{articles/030_Paper}
\includepaper{ON PRACTICAL STRATEGIES OF MELODY INDEXING FOR QUERY BY HUMMING}{Bartlomiej Stasiak}{articles/089_Paper}
\includepaper{Recurrent Neural Networks for Drum Transcription}{"Richard Vogl, Matthias Dorfer and Peter Knees"}{articles/269_Paper}
\includepaper{Integer Programming Formulation of the Problem of Generating Milton Babbitt's All-partition Arrays}{"Tsubasa Tanaka, Brian Bemman and David Meredith"}{articles/034_Paper}
\includepaper{Mining Musical Traits of Social Functions in Native American Music}{"Daniel Shanahan, Kerstin Neubarth and Darrell Conklin"}{articles/167_Paper}
\includepaper{Exploring The Latent Structure of Collaborations in Music Production}{Nazareno Andrade and Flavio Figueiredo}{articles/251_Paper}
\includepaper{Musical Note Estimation for F0 Trajectories of Singing Voices based on a Bayesian Semi-beat-synchronous HMM}{"Ryo Nishikimi, Eita Nakamura, Katsutoshi Itoyama and Kazuyoshi Yoshii"}{articles/004_Paper}
\includepaper{NOISE ROBUST MUSIC ARTIST RECOGNITION USING I-VECTOR FEATURES}{Hamid Eghbal-zadeh and Gerhard Widmer}{articles/037_Paper}
\includepaper{Piano Transcription using an Exponential Decay Model}{"Tian Cheng, Matthias Mauch, Emmanouil Benetos and Simon Dixon"}{articles/085_Paper}
\includepaper{Live Song ID: A Hashprint Approach}{"TJ Tsai, Thomas Praetzlich and Meinard Mueller"}{articles/098_Paper}
\includepaper{Nonnegative tensor factorization with frequency modulation cues for blind audio source separation}{"Elliot Creager, Noah Stein and Roland Badeau"}{articles/103_Paper}
\includepaper{On the evaluation of rhythmic and melodic descriptors}{Maria Panteli and Simon Dixon}{articles/161_Paper}
\includepaper{Analysis and classification of phonation modes in singing}{Daniel Stoller and Simon Dixon}{articles/233_Paper}
\includepaper{Beat Tracking with a Cepstroid Invariant Neural Network}{Anders Elowsson}{articles/252_Paper}
\includepaper{On drum playing technique detection in polyphonic mixtures}{Chih-Wei Wu and Alexander Lerch}{articles/268_Paper}
\includepaper{MusicDB: A Platform for Longitudinal Music Analytics}{Jeremy Hyrkas and Bill Howe}{articles/284_Paper}
\includepaper{Further Steps towards a Standard Testbed for Optical Music Recognition}{"Jan Haji_, jr., Ji_í Novotn_, Pavel Pecina and Jaroslav Pokorn_"}{articles/289_Paper}
\includepaper{Two (note) heads are better than one: pen-based multimodal interaction with music notation}{"Jorge Calvo-Zaragoza, David Rizo and Jose M. Iñesta"}{articles/006_Paper}
\includepaper{Music Structural Segmentation Across Genres with Gammatone Features}{Mi Tian and Mark B. Sandler}{articles/032_Paper}
\includepaper{Enhancing cover song identification with hierarchical rank aggregation}{"Julien Osmalskyj, Marc Van Droogenbroeck and Jean-Jacques Embrechts"}{articles/064_Paper}
\includepaper{GENRE ONTOLOGY LEARNING: COMPARING CURATED WITH CROWD-SOURCED ONTOLOGIES}{Hendrik Schreiber}{articles/074_Paper}
\includepaper{Improving voice separation by better connecting contigs}{"Nicolas Guiomard-Kagan, Mathieu Giraud, Richard Groult and Florence Levé"}{articles/129_Paper}
\includepaper{Querying XML Score Databases: XQuery is not Enough!}{"Raphael Fournier-S'niehotta, Philippe Rigaux and Nicolas Travers"}{articles/136_Paper}
\includepaper{Analysing Scattering-Based Music Content Analysis Systems: Where's the Music?}{"Francisco Rodríguez-Algarra, Bob L. Sturm and Hugo Maruri-Aguilar"}{articles/146_Paper}
\includepaper{ELUCIDATING USER BEHAVIOR IN MUSIC SERVICES THROUGH PERSONA AND GENDER}{"John Fuller, Lauren Hubener, Yea-Seul Kim and Jin Ha Lee"}{articles/205_Paper}
\includepaper{Improving Predictions of Derived Viewpoints in Multiple Viewpoints Systems}{Thomas Hedges and Geraint Wiggins}{articles/208_Paper}
\includepaper{Phrase-level Audio Segmentation of Jazz Improvisations Informed by Symbolic Data}{Jeff Gregorio and Youngmoo Kim}{articles/245_Paper}
\includepaper{Conversations with Expert Users in Music Retrieval and Research Challenges for Creative MIR}{Kristina Andersen and Peter Knees}{articles/246_Paper}
\includepaper{A comparison of melody extraction methods based on source-filter modelling}{"Juan J. Bosch, Rachel M. Bittner, Justin Salamon and Emilia Gómez"}{articles/256_Paper}
\includepaper{Large-Scale Organization of Hierarchical Music Analyses}{Phillip Kirlin}{articles/259_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Poster session 2}
\includepdf[pages=4,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Towards Template-Based Vibrato Analysis in Complex Music Signals}{"Jonathan Driedger, Stefan Balke, Sebastian Ewert and Meinard Mueller"}{articles/005_Paper}
\includepaper{"Automatic music recommendation systems: do demographic, profiling, and contextual features improve their performance?"}{Gabriel Vigliensoni and Ichiro Fujinaga}{articles/044_Paper}
\includepaper{Ensemble: A Hybrid Human-Machine System for Generating Melody Scores from Audio}{"Tim Tse, Justin Salamon, Alex Williams, Helga Jiang and Edith Law"}{articles/072_Paper}
\includepaper{An ontology for audio features}{"Alo Allik, György Fazekas and Mark Sandler"}{articles/077_Paper}
\includepaper{Predicting Missing Music Components with Bidirectional Long Short-Term Memory Neural Networks}{I-Ting Liu and Richard Randall}{articles/097_Paper}
\includepaper{Analysis of Vocal Imitations of Pitch Trajectories}{Jiajie Dai and Simon Dixon}{articles/127_Paper}
\includepaper{Whither the Design of Experiments in MIR?}{Bob L. Sturm}{articles/128_Paper}
\includepaper{Singing Voice Melody Transcription using Deep Neural Networks}{François Rigaud and Mathieu Radenen}{articles/163_Paper}
\includepaper{Musical Typicality: How Many Similar Songs Exist?}{"Tomoyasu Nakano, Daichi Mochihashi, Kazuyoshi Yoshii and Masataka Goto"}{articles/191_Paper}
\includepaper{Automatic Drum Transcription using Bi-directional Recurrent Neural Networks}{"Carl Southall, Ryan Stables and Jason Hockman"}{articles/217_Paper}
\includepaper{On the use of note onsets for improved lyrics-to-audio alignment}{"Georgi Dzhambazov, Ajay Srinivasamurthy, Sertan Sentürk and Xavier Serra"}{articles/243_Paper}
\includepaper{A higher-dimensional expansion of affective norms for English terms for music tagging}{"Michele Buccoli, Massimiliano Zanoni, György Fazekas, Augusto Sarti, Mark Sandler and Stefano Tubaro"}{articles/253_Paper}
\includepaper{Impact of Music on Decision Making in Quantitative Tasks}{"Elad Liebman, Peter Stone and Corey White"}{articles/272_Paper}
\includepaper{WiMIR: An Informetric Study on Women Authors in ISMIR}{"Xiao Hu, Kahyun Choi, Jin Ha Lee, Audrey Laplante, Yun Hao, Sally Jo Cunningham and J. Stephen Downie"}{articles/283_Paper}
\includepaper{Aligned Hierarchies: A Multi-Scale Structure-Based Representation for Music-Based Data Streams}{Katherine M. Kinnaird}{articles/020_Paper}
\includepaper{Genre specific dictionaries for harmonic/percussive source separation}{"Clément Laroche, Hélène Papadopoulos, Matthieu Kowalski and Gaël Richard"}{articles/042_Paper}
\includepaper{SPARSE CODING BASED MUSIC GENRE CLASSIFICATION USING SPECTRO_TEMPORAL MODULATIONS}{"Chih-Shan Lin, Kai-Chun Hsu and Tai-Shih Chi"}{articles/046_Paper}
\includepaper{TOWARDS EVALUATING MULTIPLE PREDOMINANT MELODY ANNOTATIONS IN JAZZ RECORDINGS}{"Stefan Balke, Jonathan Driedger, Jakob Abesser, Christian Dittmar and Meinard Mueller"}{articles/049_Paper}
\includepaper{Simultaneous Separation and Segmentation in Layered Music}{Prem Seetharaman and Bryan Pardo}{articles/057_Paper}
\includepaper{Composer Recognition based on 2D-Filtered Piano-Rolls}{"Gissel Velarde, Carlos Cancino Chacón, David Meredith, Weyde Tillman and Maaten Grachten"}{articles/063_Paper}
\includepaper{Bootstrapping a system for phoneme recognition and keyword spotting in unaccompanied singing}{Anna Kruspe}{articles/070_Paper}
\includepaper{A Methodology for Quality Assessment in Collaborative Score Libraries}{"Vincent Besson, Marco Gurrieri, Philippe Rigaux, Alice Tacaille and Virginie Thion"}{articles/076_Paper}
\includepaper{Transcribing Human Piano Performances Into Music Notation}{"Andrea Cogliati, David Temperley and Zhiyao Duan"}{articles/088_Paper}
\includepaper{Learning Invariants for Instrument Recognition}{Vincent Lostanlen and Carmine Emanuele Cella}{articles/093_Paper}
\includepaper{"Exploiting frequency, periodicity and harmonicity using advanced time-frequency concentration techniques for multipitch estimation of choir and symphony"}{"Li Su, Tsung-Ying Chuang and Yi-Hsuan Yang"}{articles/104_Paper}
\includepaper{A LATENT SPACE APPROACH TO LISTENING BEHAVIOR ANALYSIS}{"Chia-Hao Chung, Jing-Kai Lou and Homer Chen"}{articles/110_Paper}
\includepaper{Data-driven exploration of melodic structure in Hindustani music}{"Kaustuv Kanti Ganguli, Sankalp Gulati, Xavier Serra and Preeti Rao"}{articles/121_Paper}
\includepaper{learning temporal feature using deep neural network and its application to music genre classification}{Il-Young Jeong and Kyogu Lee}{articles/159_Paper}
\includepaper{Listen to Me – Don't Listen to Me: What Communities of Critics Tell Us About Music}{Ben Fields and Christophe Rhodes}{articles/173_Paper}
\includepaper{On the Potential of Simple Framewise Approaches to Piano Transcription}{"Rainer Kelz, Matthias Dorfer, Filip Korzeniowski, Sebastian Böck, Andreas Arzt and Gerhard Widmer"}{articles/179_Paper}
\includepaper{A Dataset for Automatic Practice Logging}{"R. Michael Winters, Siddharth Gururani and Alexander Lerch"}{articles/181_Paper}
\includepaper{Long-term reverberation modeling in under-determined audio source separation}{Romain Hennequin and François Rigaud}{articles/196_Paper}
\includepaper{Downbeat Detection with Conditional Random Fields and Deep Learned Features}{Simon Durand and Slim Essid}{articles/213_Paper}
\includepaper{Automatic Outlier Detection in Music Datasets}{"Yen-Cheng Lu, Chih-Wei Wu, Alexander Lerch and Chang-Tien Lu"}{articles/215_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Poster session 3}
\includepdf[pages=8,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Cross-collection evaluation for music classification tasks}{"Dmitry Bogdanov, Alastair Porter, Perfecto Herrera and Xavier Serra"}{articles/220_Paper}
\includepaper{I Said it First: Topological Analysis of Lyrical Influence Networks}{Jack Atherton and Blair Kaneshiro}{articles/221_Paper}
\includepaper{Exploring Customer Reviews for Music Genre Classification and Evolutionary Studies}{"Sergio Oramas, Luis Espinosa-Anke, Aonghus Lawlor, Horacio Saggion and Xavier Serra"}{articles/240_Paper}
\includepaper{DOWNBEAT TRACKING USING BEAT SYNCHRONOUS FEATURES WITH RECURRENT NEURAL NETWORKS}{"Florian Krebs, Sebastian Böck, Matthias Dorfer and Gerhard Widmer"}{articles/249_Paper}
\includepaper{Learning and Visualizing Music Specifications Using Pattern Graphs}{"Rafael Valle, Alexandre Donzé, Daniel J. Fremont, Ilge Akkaya, Sanjit A. Seshia and Adrian Freed"}{articles/280_Paper}
\includepaper{AVA: An Interactive System for Visual and Quantitative Analyses of Vibrato and Portamento Performance Styles}{"Luwei Yang, Elaine Chew and Khalid Z. Rajab"}{articles/314_Paper}
\includepaper{"A Hierarchical Bayesian Model of Chords, Pitches, and Spectrograms for Multipitch Analysis"}{"Yuta Ojima, Eita Nakamura, Katsutoshi Itoyama and Kazuyoshi Yoshii"}{articles/003_Paper}
\includepaper{Symbolic Meter Detection using Inner Metric Analysis}{W. Bas de Haas and Anja Volk}{articles/033_Paper}
\includepaper{CROSS TASK STUDY ON MIREX RECENT RESULTS: AN INDEX FOR EVOLUTION MEASUREMENT AND SOME STAGNATION HYPOTHESES}{"Ricardo Scholz, Geber Ramalho and Giordano Cabral"}{articles/054_Paper}
\includepaper{Towards Modeling and Decomposing Loop-Based Electronic Music}{"Patricio López-Serrano, Christian Dittmar, Jonathan Driedger and Meinard Mueller"}{articles/065_Paper}
\includepaper{Human-Interactive Optical Music Recognition}{"Liang Chen, Erik Stolterman and Christopher Raphael"}{articles/106_Paper}
\includepaper{A corpus of annotated Irish traditional dance music recordings: design and benchmark evaluations}{"Pierre Beauguitte, Bryan Duggan and John Kelleher"}{articles/148_Paper}
\includepaper{Jazz Expressive Performance Modeling}{"Helena Bantulà, Sergio Giraldo and Rafael Ramírez"}{articles/162_Paper}
\includepaper{Landmark-based Audio Fingerprinting for DJ Mix Monitoring}{"Reinhard Sonnleitner, Andreas Arzt and Gerhard Widmer"}{articles/187_Paper}
\includepaper{Mining Online Music Listening Trajectories}{"Flavio Figueiredo, Bruno Ribeiro, Jussara M. Almeida, Nazareno Andrade and Christos Faloutsos"}{articles/193_Paper}
\includepaper{Good-sounds.org: a framework to explore goodness in instrumental sounds}{"Giuseppe Bandiera, Oriol Romani Picas and Xavier Serra"}{articles/199_Paper}
\includepaper{Adaptive Frequency Neural Networks for Dynamic Pulse and Metre Perception}{"Andrew Lambert, Tillman Weyde and Newton Armstrong"}{articles/228_Paper}
\includepaper{Integration and Quality Assessment of Heterogeneous Chord Sequences using Data Fusion}{"Hendrik Vincent Koops, W. Bas de Haas, Dimitrios Bountouridis and Anja Volk"}{articles/019_Paper}
\includepaper{Can Microblogs Predict Music Charts? An Analysis of the Relationship between #Nowplaying Tweets and Music Charts}{"Eva Zangerle, Martin Pichl, Benedikt Hupfauf and Günther Specht"}{articles/039_Paper}
\includepaper{Mixtape: Direction-based navigation in large music collections}{"Joao Paulo V. Cardoso, Luciana Fujii Pontello, Pedro H. F. Holanda, Bruno Guilherme, Olga Goussevskaia and Ana Paula Couto da SIlva"}{articles/155_Paper}
\includepaper{An Evaluation Framework and Case Study for Rhythmic Concatenative Synthesis}{"Cárthach Ó Nuanáin, Perfecto Herrera and Sergi Jordà"}{articles/180_Paper}
\includepaper{Interactive Scores in Classical Music Production}{"Simon Waloschek, Axel Berndt, Benjamin W. Bohl and Aristotelis Hadjakos"}{articles/192_Paper}
\includepaper{An Analysis of Agreement in Classical Music Perception and Its Relationship to Listener Characteristics}{"Markus Schedl, Hamid Eghbal-Zadeh, Emilia Gómez and Marko Tkalcic"}{articles/260_Paper}
\includepaper{STRUCTURAL SEGMENTATION AND VISUALIZATION OF SITAR AND SAROD CONCERT AUDIO}{"Vinutha T. P., Suryanarayana Sankagiri, Kaustuv Kanti Ganguli and Preeti Rao"}{articles/151_Paper}
\includepaper{Minimax Viterbi algorithm for HMM-based Guitar fingering decision}{Gen Hori and Shigeki Sagayama}{articles/285_Paper}