### 2016-07-10 2016_Proceedings_ISMIR/keynotes.tex
updated list of paper info

%-------------------------------------------
\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Keynote Talks}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/10_Keynotes.pdf}
\thispagestyle{empty}\cleardoublepage
 %-------------------------------------------
% Keynote talk 1
%\includeabstract{Keynote 1}{Speaker 1}{external/10-1_Keynotes-1.pdf}{}
\includeabstract{Keynote 1}{Speaker 1}{}{}

\thispagestyle{empty}\cleardoublepage
% Keynote talk 2
%\includeabstract{Keynote 2}{Speaker 2}{external/10-2_Keynotes-2.pdf}{}
\includeabstract{Keynote 2}{Speaker 2}{}{}