### 2016-07-12 2016_Proceedings_ISMIR/papers.tex
updated paper list to correspond to camera ready papers

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 1: Mixed}
\includepdf[pages=2,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{SiMPle: Assessing Music Similarity Using Subsequences Joins}{Diego Silva, Chin-Chia Michael Yeh, Gustavo Batista, Eamonn Keogh}{articles/099_Paper}
\includepaper{Score-Informed Identification of Missing and Extra Notes in Piano Recordings}{Sebastian Ewert, Siying Wang, Meinard Mueller, Mark Sandler}{articles/123_Paper}
\includepaper{Feature Learning for Chord Recognition: The Deep Chroma Extractor}{Filip Korzeniowski, Gerhard Widmer}{articles/178_Paper}
\includepaper{Learning to Pinpoint Singing Voice from Weakly Labeled Examples}{Jan Schlüter}{articles/315_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 2: Rhythm}
\includepdf[pages=3,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Joint Beat and Downbeat Tracking with Recurrent Neural Networks}{Sebastian Böck, Florian Krebs, Gerhard Widmer}{articles/186_Paper}
\includepaper{Bayesian Meter Tracking on Learned Signal Representations}{Andre Holzapfel, Thomas Grill}{articles/132_Paper}
\includepaper{Tempo Estimation for Music Loops and a Simple Confidence Measure}{Frederic Font, Xavier Serra}{articles/195_Paper}
\includepaper{Brain Beats: Tempo Extraction from EEG Data}{Sebastian Stober, Thomas Praetzlich, Meinard Mueller}{articles/022_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 3: Users}
\includepdf[pages=5,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{A Plan for Sustainable MIR Evaluation}{Brian McFee, Eric Humphrey, Julián Urbano}{articles/257_Paper}
\includepaper{Go with the Flow: When Listeners Use Music as Technology}{Andrew Demetriou, Martha Larson, Cynthia C. S. Liem}{articles/068_Paper}
\includepaper{A Look at the Cloud from Both Sides Now: An Analysis of Cloud Music Service Usage}{Jin Ha Lee, Yea-Seul Kim, Chris Hubbles}{articles/087_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 4: Musicologies}
\includepdf[pages=6,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Analyzing Measure Annotations for Western Classical Music Recordings}{Christof Weiß, Vlora Arifi-Müller, Thomas Prätzlich, Rainer Kleinertz, Meinard Mueller}{articles/079_Paper}
\includepaper{Instrumental Idiom in the 16th Century: Embellishment Patterns in Arrangements of Vocal Music}{David Lewis, Tim Crawford, Daniel Müllensiefen}{articles/165_Paper}
\includepaper{The Sousta Corpus: Beat-Informed Automatic Transcription of Traditional Dance Tunes}{Andre Holzapfel, Emmanouil Benetos}{articles/156_Paper}
\includepaper{Learning a Feature Space for Similarity in World Music}{Maria Panteli, Emmanouil Benetos, Simon Dixon}{articles/166_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 5: Structure}
\includepdf[pages=7,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Systematic Exploration of Computational Music Structure Research}{Oriol Nieto, Juan Pablo Bello}{articles/043_Paper}
\includepaper{Using Priors to Improve Estimates of Music Structure}{Jordan Smith, Masataka Goto}{articles/135_Paper}
\includepaper{Music Structural Segmentation Across Genres with Gammatone Features}{Mi Tian, Mark Sandler}{articles/032_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 6: Symbolic}
\includepdf[pages=9,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Automatic Melodic Reduction Using a Supervised Probabilistic Context-Free Grammar}{Ryan Groves}{articles/274_Paper}
\includepaper{A Neural Greedy Model for Voice Separation in Symbolic Music}{Patrick Gray, Razvan Bunescu}{articles/296_Paper}
\includepaper{Towards Score Following In Sheet Music Images}{Matthias Dorfer, Andreas Arzt, Gerhard Widmer}{articles/027_Paper}
\includepaper{Extracting Ground-Truth Information from MIDI Files: A MIDIfesto}{Colin Raffel, Daniel Ellis}{articles/105_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Oral session 7: Deep-learning}
\includepdf[pages=10,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{Automatic Tagging Using Deep Convolutional Neural Networks}{Keunwoo Choi, György Fazekas, Mark Sandler}{articles/009_Paper}
\includepaper{A Hybrid Gaussian-HMM-Deep Learning Approach for Automatic Chord Estimation with Very Large Vocabulary}{Junqi Deng, Yu-Kwong Kwok}{articles/058_Paper}
\includepaper{Melody Extraction on Vocal Segments Using Multi-Column Deep Neural Networks}{Sangeun Kum, Changheun Oh, Juhan Nam}{articles/119_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Poster session 1}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{A Corpus of Annotated Irish Traditional Dance Music Recordings: Design and Benchmark Evaluations}{Pierre Beauguitte, Bryan Duggan, John Kelleher}{articles/148_Paper}
\includepaper{Adaptive Frequency Neural Networks for Dynamic Pulse and Metre Perception}{Andrew Lambert, Tillman Weyde, Newton Armstrong}{articles/228_Paper}
\includepaper{An Evaluation Framework and Case Study for Rhythmic Concatenative Synthesis}{Cárthach Ó Nuanáin, Perfecto Herrera, Sergi Jordà}{articles/180_Paper}
\includepaper{An Ontology for Audio Features}{Alo Allik, György Fazekas, Mark Sandler}{articles/077_Paper}
\includepaper{Analysis and Classification of Phonation Modes In Singing}{Daniel Stoller, Simon Dixon}{articles/233_Paper}
\includepaper{Analysis of Vocal Imitations of Pitch Trajectories}{Jiajie Dai, Simon Dixon}{articles/127_Paper}
\includepaper{Automatic Music Recommendation Systems: Do Demographic, Profiling, and Contextual Features Improve Their Performance?}{Gabriel Vigliensoni, Ichiro Fujinaga}{articles/044_Paper}
\includepaper{Automatic Outlier Detection in Music Genre Datasets}{Yen-Cheng Lu, Chih-Wei Wu, Alexander Lerch, Chang-Tien Lu}{articles/215_Paper}
\includepaper{AVA: An Interactive System for Visual and Quantitative Analyses of Vibrato and Portamento Performance Styles}{Luwei Yang, Khalid Z. Rajab, Elaine Chew}{articles/314_Paper}
\includepaper{Composer Recognition Based on 2D-Filtered Piano-Rolls}{Gissel Velarde, Tillman Weyde, Carlos Cancino Chacón, David Meredith, Maarten Grachten}{articles/063_Paper}
\includepaper{Conversations with Expert Users in Music Retrieval and Research Challenges for Creative MIR}{Kristina Andersen, Peter Knees}{articles/246_Paper}
\includepaper{Downbeat Tracking Using Beat Synchronous Features with Recurrent Neural Networks}{Florian Krebs, Sebastian Böck, Matthias Dorfer, Gerhard Widmer}{articles/249_Paper}
\includepaper{Enhancing Cover Song Identification with Hierarchical Rank Aggregation}{Julien Osmalskyj, Marc Van Droogenbroeck, Jean-Jacques Embrechts}{articles/064_Paper}
\includepaper{Ensemble: A Hybrid Human-Machine System for Generating Melody Scores from Audio}{Tim Tse, Justin Salamon, Alex Williams, Helga Jiang, Edith Law}{articles/072_Paper}
\includepaper{Exploring Customer Reviews for Music Genre Classification and Evolutionary Studies}{Sergio Oramas, Luis Espinosa-Anke, Aonghus Lawlor, Xavier Serra, Horacio Saggion}{articles/240_Paper}
\includepaper{Further Steps Towards a Standard Testbed for Optical Music Recognition}{Jan Hajič jr., Jiří Novotný, Pavel Pecina, Jaroslav Pokorný}{articles/289_Paper}
\includepaper{Improving Voice Separation by Better Connecting Contigs}{Nicolas Guiomard-Kagan, Mathieu Giraud, Richard Groult, Florence Levé}{articles/129_Paper}
\includepaper{Integer Programming Formulation of the Problem of Generating Milton Babbitt's All-Partition Arrays}{Tsubasa Tanaka, Brian Bemman, David Meredith}{articles/034_Paper}
\includepaper{Integration and Quality Assessment of Heterogeneous Chord Sequences Using Data Fusion}{Hendrik Vincent Koops, W. Bas de Haas, Dimitrios Bountouridis, Anja Volk}{articles/019_Paper}
\includepaper{Landmark-Based Audio Fingerprinting for DJ Mix Monitoring}{Reinhard Sonnleitner, Andreas Arzt, Gerhard Widmer}{articles/187_Paper}
\includepaper{Learning and Visualizing Music Specifications Using Pattern Graphs}{Rafael Valle, Daniel J. Fremont, Ilge Akkaya, Alexandre Donzé, Adrian Freed, Sanjit A. Seshia}{articles/280_Paper}
\includepaper{Listen To Me – Don't Listen To Me: What Communities of Critics Tell Us About Music}{Ben Fields, Christophe Rhodes}{articles/173_Paper}
\includepaper{Long-Term Reverberation Modeling for Under-Determined Audio Source Separation with Application to Vocal Melody Extraction}{Romain Hennequin, François Rigaud}{articles/196_Paper}
\includepaper{Nonnegative Tensor Factorization with Frequency Modulation Cues for Blind Audio Source Separation}{Elliot Creager, Noah Stein, Roland Badeau, Philippe Depalle}{articles/103_Paper}
\includepaper{On Drum Playing Technique Detection in Polyphonic Mixtures}{Chih-Wei Wu, Alexander Lerch}{articles/268_Paper}
\includepaper{Predicting Missing Music Components with Bidirectional Long Short-Term Memory Neural Networks}{I-Ting Liu, Richard Randall}{articles/097_Paper}
\includepaper{Structural Segmentation and Visualization of Sitar and Sarod Concert Audio}{Vinutha T. P., Suryanarayana Sankagiri, Kaustuv Kanti Ganguli, Preeti Rao}{articles/151_Paper}
\includepaper{Template-Based Vibrato Analysis in Complex Music Signals}{Jonathan Driedger, Stefan Balke, Sebastian Ewert, Meinard Mueller}{articles/005_Paper}
\includepaper{Towards Evaluating Multiple Predominant Melody Annotations in Jazz Recordings}{Stefan Balke, Jonathan Driedger, Jakob Abesser, Christian Dittmar, Meinard Mueller}{articles/049_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Poster session 2}
\includepdf[pages=4,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{A Hierarchical Bayesian Model of Chords, Pitches, and Spectrograms for Multipitch Analysis}{Yuta Ojima, Eita Nakamura, Katsutoshi Itoyama, Kazuyoshi Yoshii}{articles/003_Paper}
\includepaper{A Higher-Dimensional Expansion of Affective Norms for English Terms for Music Tagging}{Michele Buccoli, Massimiliano Zanoni, György Fazekas, Augusto Sarti, Mark Sandler}{articles/253_Paper}
\includepaper{A Latent Representation of Users, Sessions, and Songs for Listening Behavior Analysis}{Chia-Hao Chung, Jing-Kai Lou, Homer Chen}{articles/110_Paper}
\includepaper{A Methodology for Quality Assessment in Collaborative Score Libraries}{Vincent Besson, Marco Gurrieri, Philippe Rigaux, Alice Tacaille, Virginie Thion}{articles/076_Paper}
\includepaper{Aligned Hierarchies: A Multi-Scale Structure-Based Representation for Music-Based Data Streams}{Katherine M. Kinnaird}{articles/020_Paper}
\includepaper{Analysing Scattering-Based Music Content Analysis Systems: Where's the Music?}{Francisco Rodríguez-Algarra, Bob L. Sturm, Hugo Maruri-Aguilar}{articles/146_Paper}
\includepaper{Beat Tracking with a Cepstroid Invariant Neural Network}{Anders Elowsson}{articles/252_Paper}
\includepaper{Bootstrapping a System for Phoneme Recognition and Keyword Spotting in Unaccompanied Singing}{Anna Kruspe}{articles/070_Paper}
\includepaper{Can Microblogs Predict Music Charts? An Analysis of the Relationship Between \#Nowplaying Tweets and Music Charts}{Eva Zangerle, Martin Pichl, Benedikt Hupfauf, Günther Specht}{articles/039_Paper}
\includepaper{Cross Task Study on MIREX Recent Results: An Index for Evolution Measurement and Some Stagnation Hypotheses}{Ricardo Scholz, Geber Ramalho, Giordano Cabral}{articles/054_Paper}
\includepaper{Cross-Collection Evaluation for Music Classification Tasks}{Dmitry Bogdanov, Alastair Porter, Perfecto Herrera, Xavier Serra}{articles/220_Paper}
\includepaper{Downbeat Detection with Conditional Random Fields and Deep Learned Features}{Simon Durand, Slim Essid}{articles/213_Paper}
\includepaper{Exploiting Frequency, Periodicity and Harmonicity Using Advanced Time-Frequency Concentration Techniques for Multipitch Estimation of Choir and Symphony}{Li Su, Tsung-Ying Chuang, Yi-Hsuan Yang}{articles/104_Paper}
\includepaper{Genre Ontology Learning: Comparing Curated with Crowd-Sourced Ontologies}{Hendrik Schreiber}{articles/074_Paper}
\includepaper{Genre Specific Dictionaries for Harmonic/Percussive Source Separation}{Clément Laroche, Hélène Papadopoulos, Matthieu Kowalski, Gaël Richard}{articles/042_Paper}
\includepaper{Good-sounds.org: A Framework to Explore Goodness in Instrumental Sounds}{Giuseppe Bandiera, Oriol Romani Picas, Xavier Serra}{articles/199_Paper}
\includepaper{Improving Predictions of Derived Viewpoints in Multiple Viewpoints Systems}{Thomas Hedges, Geraint Wiggins}{articles/208_Paper}
\includepaper{Known Artist Live Song ID: A Hashprint Approach}{TJ Tsai, Thomas Praetzlich, Meinard Mueller}{articles/098_Paper}
\includepaper{Learning Temporal Features Using a Deep Neural Network and its Application to Music Genre Classification}{Il-Young Jeong, Kyogu Lee}{articles/159_Paper}
\includepaper{Meter Detection in Symbolic Music Using Inner Metric Analysis}{W. Bas de Haas, Anja Volk}{articles/033_Paper}
\includepaper{Minimax Viterbi Algorithm for HMM-Based Guitar Fingering Decision}{Gen Hori, Shigeki Sagayama}{articles/285_Paper}
\includepaper{Mixtape: Direction-Based Navigation in Large Media Collections}{Joao Paulo V. Cardoso, Luciana Fujii Pontello, Pedro H. F. Holanda, Bruno Guilherme, Olga Goussevskaia, Ana Paula Couto da Silva}{articles/155_Paper}
\includepaper{Musical Note Estimation for F0 Trajectories of Singing Voices Based on a Bayesian Semi-Beat-Synchronous HMM}{Ryo Nishikimi, Eita Nakamura, Katsutoshi Itoyama, Kazuyoshi Yoshii}{articles/004_Paper}
\includepaper{On the Evaluation of Rhythmic and Melodic Descriptors for Music Similarity}{Maria Panteli, Simon Dixon}{articles/161_Paper}
\includepaper{On the Potential of Simple Framewise Approaches to Piano Transcription}{Rainer Kelz, Matthias Dorfer, Filip Korzeniowski, Sebastian Böck, Andreas Arzt, Gerhard Widmer}{articles/179_Paper}
\includepaper{Phrase-Level Audio Segmentation of Jazz Improvisations Informed by Symbolic Data}{Jeff Gregorio, Youngmoo Kim}{articles/245_Paper}
\includepaper{Revisiting Priorities: Improving MIR Evaluation Practices}{Bob L. Sturm}{articles/128_Paper}
\includepaper{Simultaneous Separation and Segmentation in Layered Music}{Prem Seetharaman, Bryan Pardo}{articles/057_Paper}
\includepaper{Towards Modeling and Decomposing Loop-Based Electronic Music}{Patricio López-Serrano, Christian Dittmar, Jonathan Driedger, Meinard Mueller}{articles/065_Paper}
\includepaper{Two (Note) Heads Are Better Than One: Pen-Based Multimodal Interaction with Music Scores}{Jorge Calvo-Zaragoza, David Rizo, Jose M. Iñesta}{articles/006_Paper}

\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Poster session 3}
\includepdf[pages=8,pagecommand=\thispagestyle{empty}]{external/12_Sessions.pdf}%
\thispagestyle{empty}\cleardoublepage

\includepaper{A Comparison of Melody Extraction Methods Based on Source-Filter Modelling}{Juan J. Bosch, Rachel M. Bittner, Justin Salamon, Emilia Gómez}{articles/256_Paper}
\includepaper{An Analysis of Agreement in Classical Music Perception and its Relationship to Listener Characteristics}{Markus Schedl, Hamid Eghbal-Zadeh, Emilia Gómez, Marko Tkalcic}{articles/260_Paper}
\includepaper{An Attack/Decay Model for Piano Transcription}{Tian Cheng, Matthias Mauch, Emmanouil Benetos, Simon Dixon}{articles/085_Paper}
\includepaper{Automatic Drum Transcription Using Bi-Directional Recurrent Neural Networks}{Carl Southall, Ryan Stables, Jason Hockman}{articles/217_Paper}
\includepaper{Automatic Practice Logging: Introduction, Dataset \& Preliminary Study}{R. Michael Winters, Siddharth Gururani, Alexander Lerch}{articles/181_Paper}
\includepaper{Data-Driven Exploration of Melodic Structure in Hindustani Music}{Kaustuv Kanti Ganguli, Sankalp Gulati, Xavier Serra, Preeti Rao}{articles/121_Paper}
\includepaper{Deep Convolutional Networks on the Pitch Spiral For Music Instrument Recognition}{Vincent Lostanlen, Carmine Emanuele Cella}{articles/093_Paper}
\includepaper{DTV-Based Melody Cutting for DTW-Based Melody Search and Indexing in QbH Systems}{Bartlomiej Stasiak}{articles/089_Paper}
\includepaper{Elucidating User Behavior in Music Services Through Persona and Gender}{John Fuller, Lauren Hubener, Yea-Seul Kim, Jin Ha Lee}{articles/205_Paper}
\includepaper{Exploring the Latent Structure of Collaborations in Music Recordings: A Case Study in Jazz}{Nazareno Andrade, Flavio Figueiredo}{articles/251_Paper}
\includepaper{Global Properties of Expert and Algorithmic Hierarchical Music Analyses}{Phillip Kirlin}{articles/259_Paper}
\includepaper{Human-Interactive Optical Music Recognition}{Liang Chen, Erik Stolterman, Christopher Raphael}{articles/106_Paper}
\includepaper{I Said it First: Topological Analysis of Lyrical Influence Networks}{Jack Atherton, Blair Kaneshiro}{articles/221_Paper}
\includepaper{Impact of Music on Decision Making in Quantitative Tasks}{Elad Liebman, Peter Stone, Corey White}{articles/272_Paper}
\includepaper{Interactive Scores in Classical Music Production}{Simon Waloschek, Axel Berndt, Benjamin W. Bohl, Aristotelis Hadjakos}{articles/192_Paper}
\includepaper{Jazz Ensemble Expressive Performance Modeling}{Helena Bantulà, Sergio Giraldo, Rafael Ramírez}{articles/162_Paper}
\includepaper{Mining Musical Traits of Social Functions in Native American Music}{Daniel Shanahan, Kerstin Neubarth, Darrell Conklin}{articles/167_Paper}
\includepaper{Mining Online Music Listening Trajectories}{Flavio Figueiredo, Bruno Ribeiro, Christos Faloutsos, Nazareno Andrade, Jussara M. Almeida}{articles/193_Paper}
\includepaper{Musical Typicality: How Many Similar Songs Exist?}{Tomoyasu Nakano, Daichi Mochihashi, Kazuyoshi Yoshii, Masataka Goto}{articles/191_Paper}
\includepaper{MusicDB: A Platform for Longitudinal Music Analytics}{Jeremy Hyrkas, Bill Howe}{articles/284_Paper}
\includepaper{Noise Robust Music Artist Recognition Using I-Vector Features}{Hamid Eghbal-zadeh, Gerhard Widmer}{articles/037_Paper}
\includepaper{On the Use of Note Onsets for Improved Lyrics-To-Audio Alignment in Turkish Makam Music}{Georgi Dzhambazov, Ajay Srinivasamurthy, Sertan Sentürk, Xavier Serra}{articles/243_Paper}
\includepaper{Querying XML Score Databases: XQuery is not Enough!}{Raphael Fournier-S'niehotta, Philippe Rigaux, Nicolas Travers}{articles/136_Paper}
\includepaper{Recurrent Neural Networks for Drum Transcription}{Richard Vogl, Matthias Dorfer, Peter Knees}{articles/269_Paper}
\includepaper{Singing Voice Melody Transcription Using Deep Neural Networks}{François Rigaud, Mathieu Radenen}{articles/163_Paper}
\includepaper{Sparse Coding Based Music Genre Classification Using Spectro-Temporal Modulations}{Kai-Chun Hsu, Chih-Shan Lin, Tai-Shih Chi}{articles/046_Paper}
\includepaper{Time-Delayed Melody Surfaces for Rāga Recognition}{Sankalp Gulati, Joan Serrà, Kaustuv Kanti Ganguli, Sertan Şentürk, Xavier Serra}{articles/030_Paper}
\includepaper{Transcribing Human Piano Performances into Music Notation}{Andrea Cogliati, David Temperley, Zhiyao Duan}{articles/088_Paper}
\includepaper{WIMIR an Informetric Study on Women Authors in ISMIR}{Xiao Hu, Kahyun Choi, Jin Ha Lee, Audrey Laplante, Yun Hao, Sally Jo Cunningham, J. Stephen Downie}{articles/283_Paper}