### 2016-07-13 2016_Proceedings_ISMIR/keynotes.tex
big update of pre-defined pdfs from google docs, everything in place now except cover, need to finalize preface

%-------------------------------------------
\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Keynote Talks}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/10_Keynotes.pdf}
\thispagestyle{empty}\cleardoublepage
 %-------------------------------------------

% Keynote talk 1
\addcontentsline{toc}{subsection}{The New York Philharmonic Leon Levy Digital Archives: An Integrated View of Performance Since 1842\\\textit{Barbara Haws}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/10-1_Keynotes-1.pdf}%

\thispagestyle{empty}\cleardoublepage
% Keynote talk 2
\addcontentsline{toc}{subsection}{Is Machine Learning Enough?\\\textit{Beth Logan, PhD}}
\includepdf[pagecommand=\thispagestyle{fancy}]{external/10-2_Keynotes-2.pdf}%