### 2016-07-13 2016_Proceedings_ISMIR/tutorials.tex

%-------------------------------------------
\thispagestyle{empty}\cleardoublepage
\addcontentsline{toc}{section}{Tutorials}
\includepdf[pages=1,pagecommand=\thispagestyle{empty}]{external/11_Tutorials.pdf}%
\thispagestyle{empty}\cleardoublepage
 %-------------------------------------------

\newcommand{\includetut}[3]{\addcontentsline{toc}{subsection}{#1\\\textit{#2}}
\includepdf[pages=#3, pagecommand=\thispagestyle{fancy}]{external/11_TutorialsAll.pdf}}

% Tutorial 1
\includetut{Jazz solo analysis between music information
  retrieval, music psychology, and jazz research}{Jakob Abe{\ss}er,
  Klaus Frieler, Wolf-Georg Zaddach}{1-2}

% Tutorial 2
\includetut{Music Information Retrieval: Overview, Recent
  Developments and Future Challenges}{Emilia G\'omez, Markus Schedl,
  Xavier Serra, Xiao Hu}{3}

% Tutorial 3
\includetut{Why is studio production interesting?}{Emmanuel
Deruty, Fran\c{c}ois Pachet}{4-5}

% Tutorial 4
\includetut{Introduction to EEG Decoding for Music Information Retrieval
Research}{Sebastian Stober, Blair Kaneshiro}{6}

% Tutorial 5
\includetut{Natural Language Processing for MIR}{Sergio Oramas,
  Luis Espinosa-Anke, Shuo Zhang, Horacio Saggion}{7}

% Tutorial 6
\includetut{Why Hip-Hop is interesting}{Jan Van Balen, Ethan
  Hein, Dan Brown}{8-9}